---
title: "Fig 4"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---


```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(RColorBrewer)
library(grid)
library(gridExtra)
library(here)
library(ggstance)
library(readxl)
library(devtools)
devtools::load_all(here("code","liver.tools"))
```
Read in Noushin's data for TCGA figure

```{r data}
fig.data <- readRDS(here('data', 'tcga', 'data.rds'))
```

## TCGA copy number

TCGA figure panel -- problems, chromosome widths, making the panel match size of one facet in z-score panel

```{r plot, fig.height= 23, fig.width=8,dev=c("png", "pdf")}
setDT(fig.data)
fig.data[,bin:=as.factor(rev(bin))][]
fig.data$disease<-"TCGA Ovarian Cancers\n(n = 597)"
plot <- fig.data %>%
    mutate(disease="TCGA Ovarian Cancers\n(n = 597)") %>%
    ggplot(aes(x=value, y=bin, color=change, fill=change)) +
    facet_grid(arm~disease, scales = 'free_y') + ##,
    ##           switch = 'y')+#, space = 'free')+
    ##facet_wrap(~arm, ncol=1, scales="free_y") + ##
    geom_col() +
    scale_x_continuous() +
    #scale_fill_brewer(palette = 'Set1')+
    #scale_color_brewer(palette = 'Set1')+
    scale_color_manual(values=c("red3","#18468B"))+
    scale_fill_manual(values=c("red3","#18468B"))+
    theme_minimal(base_size=24) +
    theme(strip.text.y = element_text(angle = 0,
                                      hjust = 0.5,
                                      vjust = 0.5,
                                      size=20),
          strip.text.x=element_text(size=20),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          panel.grid =  element_blank(),
          legend.title = element_blank(),
          legend.position="bottom",
          axis.title.x = element_blank(),
          strip.background = element_blank()) + ##removing label
    #xlab("Proportion of\ncases with CNV") +
    ylab("Chromosome Arm")
```

## Ovarian study z-scores

The z-score's with our data panel. Problems -- chromosome width.

```{r z_scores}
meta<-fread(here("data","metadata.csv"))
meta<-meta %>% select(-V1) %>% mutate(classifier_type2=if_else(id=="CGPLOV161P","Benign",classifier_type2))

data<-fread(here("data","screening_train_data.csv")) %>% select(-type)
data2<-fread(here("data","diag_train_data.csv")) %>% filter(type=="healthy") %>% select(-type)
data<-rbind(data,data2)

data<-inner_join(data,meta %>% select(id,classifier_type),by="id") %>% rename(type=classifier_type)



features<-data

cols <- brewer.pal(7,"Set1")
cols[3] <- cols[4]

features <- features %>% select(id,starts_with("z"))
features <- inner_join(features, meta %>%
                                     select(id,`classifier_type2`),
                           by=c("id"="id"))




datx<-read_csv(here("data/scores","All_Scores_CV.csv"))
datx<-datx %>% filter(training_set_include_screening=="yes"|training_set_include_diag=="yes")
datx<-datx %>% select(id,classifier_type,`Screening ZEUS`,`Diagnostic ZEUS`,subtype_simple,stage)
datx<-datx %>% gather(key=model,value=score,-stage,-subtype_simple,-classifier_type,-id) %>% filter(!is.na(score))

features<-features %>% filter(id %in% datx$id)

features <- features %>%
        gather(key="Arm", value="Zscore", starts_with("z"))
    features$Arm <- sapply( str_split(features$Arm,"_"), "[", 2 )

    chrlevels <- c("1p","1q","2p","2q","3p","3q","4p","4q","5p",
                   "5q","6p","6q","7p","7q","8p","8q","9p","9q",
                   "10p","10q","11p","11q","12p","12q",
                   "13q","14q","15q","16p","16q","17p","17q",
                   "18p","18q","19p","19q","20p","20q","21q","22q")
    features$Arm <- factor(features$Arm,
                           levels = chrlevels)
    features <- features %>% mutate(colors=if_else(Zscore<=0,"Negative","Positive"))
    ##features$colors<-factor(features$colors,levels=c("Positive","Negative"))
    ##Z-score for two-tailed p-value .05 w/ bonferroni correction for 39 arms
    features <- features %>%
        mutate(colors=if_else(Zscore>= -3.22 & Zscore <= 3.22 ,
                              "Neutral", features$colors))
    ##features<- features %>% mutate(transp=if_else(colors=="Neutral",.9,.2))
    ##features<- features %>% mutate(transp=if_else(colors=="Neutral",.05,.99))
    features <- features %>% mutate(transp=if_else(colors=="Neutral",1,1))
    ##features<-features %>% mutate(colors=if_else(Zscore<=0,"Negative","Positive"))
    features$colors<-factor(features$colors,levels=c("Positive","Negative","Neutral"))
    features$y=1
    ##features$log<-log10(abs(features$Zscore))
    ##features <- features %>% mutate(log=if_else(Zscore<=0,-1*log,log))
    features <- features %>% mutate(Zscore=if_else(Zscore<=-150,-150,Zscore))
    features <- features %>% mutate(Zscore=if_else(Zscore>150,150,Zscore))
    features <- features %>%
        mutate(`type`=if_else(`classifier_type2`=="cancer","Ovarian Cancer",classifier_type2))
      features <- features %>%
        mutate(`type`=if_else(`classifier_type2`=="non-cancer","Non-cancer",type))

    features$type <- factor(features$`type`,
                                  levels=c("Ovarian Cancer","Benign","Non-cancer"))
    features2  <- features %>%
        rename(z=Zscore) %>%
        mutate(root2=nroot(z, 2)) %>%
        mutate(root3=nroot(z, 3))


features2 <- features2  %>%
    select(id, type, Arm,
           z, root2, root3, y, colors, transp)
###YOU ARE HERE AFTER LUNCH!!!

b<-features2 %>% filter(type=="Benign")
#b2<-meta %>% select(-V1) %>% filter(classifier_type2=="Benign") %>% filter(training_set_include_diag=="yes")
ns <- features2 %>%
    group_by(type) %>%
    summarize(n=length(unique(id))) %>%
    mutate(label=paste0(type, "\n(n = ", n, ")"))
ns <- setNames(ns$label, ns$type)
cols[2]<-"#18468B"
cols[1]<-"red3"

zplot_cancer <- features2 %>%
    ggplot(aes(x=y, y=z, color=colors, alpha=transp)) +
    geom_point(size=5) +
    facet_grid(Arm ~ type, scales = 'free_y',labeller=labeller(type=ns)) +
    coord_flip() +
    theme_minimal(base_size=24) +
    theme(strip.text.y.left = element_text(angle = 0,
                                           hjust = 0.5,
                                           vjust = 0.5),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),

          panel.grid =  element_blank(),
          plot.title = element_blank(),
          legend.title = element_blank(),
          axis.title.y=element_blank(),
          legend.position = "bottom",
          strip.text.y = element_blank(),
          axis.title.x = element_blank()) + #####removing label
    scale_y_continuous(limits = c(-150,150)) +
    scale_alpha_identity() +
    scale_fill_manual(values = cols) +
    scale_color_manual(values = cols) #+
    #ylab("Z-Score")
```

Adding together the TCGA and z-score panel -- problems -- chromosome width, size of TCGA panel, not yet stacked with feature importance panel.

```{r cowplot, eval=FALSE}
#pdf("test.pdf")
cowplot::plot_grid(plot,zplot_cancer,nrow=1,align="v",axis="l")
#dev.off() #
```

## Feature importance

Feature importance panel
```{r heatmap}

results<-readRDS(here("output","feature_importance.Rmd",
                     "loadings.rds"))
pd <- results[["pd"]]
sc <- results[["sc"]]
#---------------- now make plots -----------------###
                                        # visualize feature by PC heatmap ##
pc.labels <- c("PC1\n", "PC2\n", "PC3\n") %>%
    setNames(1:3)
heatmap <- pd %>% filter(variable != "PC1")%>%
    ggplot(aes(x = pc.id, y = bin.id)) +
    facet_grid(arm ~ pc.id, scales = 'free',
               switch = 'y', space = 'free',
               labeller=labeller(pc.id=pc.labels))+
    geom_tile(aes(fill = value)) +
    scale_fill_distiller(palette = 'RdBu') +
    scale_y_continuous(trans = 'reverse') +
    theme_minimal(base_size=24) +
    labs(x = 'Principal components\nof fragmentation profiles', y = '', fill = 'Value') +
    theme(axis.text.x=element_text(color="white"),
          strip.text.y=element_blank(),
          strip.text.x=element_text(size=26),
          axis.text.y = element_text(color="white"),
          axis.title.y=element_text(color="white"),
          axis.ticks = element_blank(),
          panel.grid =  element_blank(),
          legend.position = 'none') +
    ylab("Something")
```

```{r top_annotation}
ta <- subset(sc, grepl('PC', features))
ta$pc.id <- gsub('PC', '', gsub(' ','',ta$features))
ta$sign.value = factor(ta$sign.value, levels = c(-1, 1))
ta <- ta[,c('pc.id', 'abs.value', 'sign.value')]
#ta <- rbind(ta, data.frame(pc.id = setdiff(seq(3), ta$pc.id), abs.value =0, sign.value = NA))
ta$pc.id <- factor(ta$pc.id, levels = ta$pc.id)
ta$x = 1
ta$arm = 1
ta


cols <- c('#0571b0', '#ca0020', '#999999')
names(cols) <- c('-1', '1', NA)
top.annot <- ta %>%
    ggplot(aes(x = pc.id, y = abs.value, colour = sign.value)) +
    facet_grid(arm ~ pc.id, scales = 'free',
                   switch = 'y', space = 'free') +
    geom_point(size = 5) +
    geom_segment(aes(x = pc.id,
                     xend = pc.id,
                     y = abs.value, yend = 0),
                 size=2) +
    theme_minimal(base_size=24) +
    ylab("Log odds ratio") +
    xlab("") +
    theme(axis.text.x=element_text(color="white"),
          strip.text.y=element_blank(),
          axis.line.x=element_line(color="gray"),
          axis.ticks.x=element_line(color="gray"),
          axis.ticks = element_blank(),
          panel.grid =  element_blank(),
          legend.position = 'none') +
##    theme(panel.grid=element_blank(),
##          strip.text.y = element_blank(),

##          strip.text.x = element_blank(),
##          axis.text.x = element_blank(),
##          axis.ticks = element_blank(),
##          panel.border = element_rect(color = "white",
##                                      fill = NA, size = 0.2),
##          legend.position = 'none') +
    scale_color_manual(values = cols, na.value = '#999999')
```

```{r side_annotation}
bins5mb <- read_csv(here("data","long_bins_hiseq.csv"),
                    show_col_types=FALSE) 
bins5mb<-bins5mb %>% filter(id %in% features$id)
bins5mb<-bins5mb %>% select(-`...1`)
sa <- side_annotation(pd, sc, bins5mb)


#sa2 <- results2[["sc"]]
#sa2<-sa2 %>% filter(! features %in% c("PC 1","PC 2", "PC 3","TF_pc_1")) 
#pd2 <- results2[["pd"]]
#sa_temp <- side_annotation(pd2, sc2, bins5mb)$arm.imp
#x <- strsplit(as.character(sa2$features), " " )

#sa2$arm<-sapply( x, "[", 2 )
#sa2<-left_join(sa_temp %>% select(-abs.value),sa2 %>% select(arm,abs.value,scaled_coefs),by="arm")
#sa2 <- sa2 %>% mutate(abs.value=if_else(is.na(abs.value),0,abs.value))
#sa2<-sa2 %>% mutate(sign.value=sign(scaled_coefs))
#sa2 <- sa2 %>% select(-scaled_coefs)

bl <- sa[["bl"]]
arm.imp <- sa[["arm.imp"]]
#arm.imp$model<-"DELFI"
#sa2$model<-"DELFI+TF"
#arm.imp<-rbind(arm.imp,sa2)
right.annot <- bl %>%
    ggplot(aes(x = x, y = bin.id)) +
    geom_point(color = 'white') +
    facet_grid(arm ~ pc.id,
               scales = 'free',
               switch = 'y', space = 'free') +
    theme_minimal(base_size=24) +
    geom_point(data = arm.imp,
               aes(x = abs.value,
                   y = bin.id, color = sign.value),
               size=5) +
    geom_segment(data = arm.imp,
                 aes(x = abs.value, xend = 0,
                     y = bin.id, yend = bin.id,
                     color = sign.value),
                 size=2) +
    scale_color_manual(values = cols, na.value = '#999999') +
    scale_y_continuous(breaks = c(0,2,4,6)) +
    labs(x = 'Log odds\nratio', y = '', color = '') +
    theme(strip.text.x = element_blank(),
          strip.text.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          panel.border = element_blank(),
          panel.grid=element_blank(),
          legend.position = 'none')



```

```{r extra_annot}

cols <- c('#0571b0', '#ca0020', '#999999')
names(cols) <- c('-1', '1', NA)
p<-sc %>% filter(features %in% c("ca125","he4"))
p$feature.type<-"Protein"
extra.annot <- p %>%
    ggplot(aes(x = features, y = abs.value, colour = sign.value)) +
    geom_point(size = 5) +
    geom_segment(aes(x = features,
                     xend = features,
                     y = abs.value, yend = 0),
                 size=2) +
    theme_minimal(base_size=24) +
    ylab("Log odds ratio") +
    xlab("") +
    theme(axis.text.x=element_text(color="black"),
          strip.text.y=element_blank(),
          axis.line.x=element_line(color="gray"),
          axis.ticks.x=element_line(color="gray"),
          axis.ticks = element_blank(),
          panel.grid =  element_blank(),
          legend.position = 'none') +
##    theme(panel.grid=element_blank(),
##          strip.text.y = element_blank(),

##          strip.text.x = element_blank(),
##          axis.text.x = element_blank(),
##          axis.ticks = element_blank(),
##          panel.border = element_rect(color = "white",
##                                      fill = NA, size = 0.2),
##          legend.position = 'none') +
    scale_color_manual(values = cols, na.value = '#999999')


extra.annot
```


## Fig 4

Assembling the 3 pieces of the feature importance figure into one panel

```{r heights, fig.height=23, fig.width=10, dev=c("png", "pdf")}
hmap <- ggplotGrob(heatmap)
z.coef <- ggplotGrob(right.annot)
pc.coef <- ggplotGrob(top.annot)
h <- hmap$heights
## I think these widths correspond to the panels
ix <- seq(from=8, by=2, length.out=39)
## make logical
is_arm <- seq_along(h) %in% ix
h2 <- h[is_arm]
size <- map_dbl(h2, function(x) unlist(x)[[1]])
## The widths less than 100 might be too small
is2small <- size < 12
h2[is2small] <- unit(12, "null")
h[is_arm] <- h2
z.coef$heights = h
hmap$heights <- h
hmap$widths <- pc.coef$widths
#saveRDS(h, file="../data/arm_heights.rds")

extra.coef <- ggplotGrob(extra.annot)

```


```{r fig4, fig.height=25, fig.width=28, dev=c("png", "pdf","svg")}
#tf<-ggplotGrob(tf_annotation)
#tf$widths<-z.coef$widths
#tf$heights<-z.coef$heights
tcga.legend <- cowplot::get_legend(plot)
p2 <- plot +
    guides(fill="none", color="none") +
    ylab("")
g1 <- ggplotGrob(p2)
g1$heights <- h

zscore.legend <- cowplot::get_legend(zplot_cancer)
p3 <- zplot_cancer +
    guides(color="none", fill="none")
g2 <- ggplotGrob(p3)
g2$heights <- h
gl <- list(g1,
           g2,
           pc.coef,
           hmap,
           z.coef,
          nullGrob(),extra.coef)
m <- rbind(c(6, 6, 3, 7),
           c(1, 2, 4, 5))
grid.arrange(grobs=gl,
             layout_matrix=m,
             heights=c(1, 7),
             widths=c(2.7, 4, 4, 2.7),
             padding=unit(0, "lines")) ##
```

