---
title: "Predictive distributions of performance in future studies"
---

# Goals

In a future study of 100,000 participants, what is our predictive distribution for the following:
- number individuals screened
- liver cancers detected by delfi
- liver cancers detected by ultrasound (US) + alpha fetal protein (AFP)
- false positives leading to unnecessary follow-up procedures by these approaches
- individuals with liver cancer not identified ( false negatives )

The size of the screening study is determined by the number of samples collected in a finite period of time.  Studies with lower adherence to surveillance guidelines have fewer participants.


# Best guesses

```{r packages, message=FALSE, echo=FALSE}
library(here)
library(devtools)
devtools::load_all(here("code", "liver.tools"))
library(grid)
library(gridExtra)
library(gtools)
library(tidyverse)
library(mvtnorm)
library(magrittr)
library(ggplot2)
library(epiR)
library(readxl)
library(data.table)
library(patchwork)
library(cowplot)
tools <- here("data", "simulations_metrics.csv") %>%
    fread() %>% filter(tool %in% c("CA125_30_ukctocs","CA125_30_jhu","HE4_70_jhu","HE4_70_Jacob","DELFI-Pro"))

set.seed(1949694)

```

## We have a point estimate with a confidence interval -- binomial for most things, beta for the adherence stuff based on the liver paper

```{r betas}
find_params <- function(i, dat){
    dat <- dat[i, ]
    params <- epi.betabuster(mode=dat$mode,
                             conf=0.975,
                             ##greaterthan=TRUE,
                             imsure="greater than",
                             x=dat$lower,
                             max.shape1=500,
                             conf.level=0.95)
    dat$shape1 <- params$shape1
    dat$shape2 <- params$shape2
    dat
}
params <- seq_len(nrow(tools)) %>%
    map_dfr(find_params, dat=tools)
```


```{r performance}
simulate_performance <- function(i, params, N=1000){
    p <- params[i, ]
    x <- rbeta(N, shape1=p$shape1, shape2=p$shape2)
    x
}
simvalues <- seq_len(nrow(tools)) %>%
    map(simulate_performance, params=params)
params$simulation <- simvalues
params2 <- params %>%
    filter(metric != "adherence") %>%
    pivot_wider(id_cols=c("tool", "population"),
                names_from="metric",
                values_from="simulation") %>%
    unnest(c("sensitivity", "specificity"))
adhere <- params %>%
    filter(metric=="adherence") %>%
    unnest(simulation) %>%
    select(tool, population, simulation) %>%
    set_colnames(c("tool", "population", "adherence")) %>%
    nest(adherence=adherence)
params3 <- params2 %>%
    nest(performance=c(sensitivity, specificity))
params4 <- left_join(params3, adhere,
                     by=c("tool", "population"))
panelA.data <- params2 #%>%
   # mutate(population=factor(population,
                             #rev(c("High risk", "Early stage"))))
```

```{r panelA}
panelA.data<-panelA.data #%>% filter(tool %in% c("DELFI", "US + AFP")) 
#panelA.data$tool<-factor(panelA.data$tool,levels=c("US + AFP","DELFI"))
pt_esti<-tools %>% select(tool,mode,metric) %>% filter(metric %in% c("sensitivity","specificity"))%>%spread(key=metric,value=mode)
panelA.data$tool<-factor(panelA.data$tool,levels=c("HE4_70_Jacob","HE4_70_jhu","CA125_30_ukctocs","CA125_30_jhu","DELFI-Pro"))
panelA <- panelA.data %>% 
    ggplot(aes(specificity, sensitivity)) +
    theme_classic(base_size=20) +
    theme(panel.grid=element_blank()) +
    geom_density2d(aes(color=tool),bins=25, size=.5) +
    scale_y_continuous(expand=c(0, 0), limits=c(0, 1)) +
    scale_x_continuous(expand=c(0, 0), limits=c(0, 1)) +
    geom_point(data=pt_esti,aes(specificity, sensitivity), size=3) +
    ##xlim(c(0.15, 1)) +
    ##ylim(c(0.15, 1)) +
    ##facet_wrap(~population, ncol=1) +
    theme(strip.background=element_blank(),
          legend.position=c(0.2, 0.25)) +
    guides(color=guide_legend(title="",
                              override.aes=list(size=1.5))) +
    xlab("Specificity") +
    ylab("Sensitivity")+theme(legend.position="left")+
  scale_color_manual(values=c("steelblue","steelblue1","lightpink1","deeppink","darkmagenta"),
                     labels=c("HE4 - 70 pM - Jacob et al., 2011","HE4 - 70 pM - JHU","CA125 - 30 U/mL - Menon et al., 2015","CA125 - 30 U/mL - JHU", "DELFI-Pro - >99% Specificity"))+coord_fixed()
panelA
#panelA <- ggplotGrob(panelA)
```

```{r panelA2, fig.width=4, fig.height=7, dev=c("png", "pdf")}
#panelA.data$tool<-factor(panelA.data$tool,levels=c("US + AFP","DELFI"))
panelA2 <- panelA.data %>% 
    ggplot(aes(specificity, sensitivity)) +
    theme(panel.grid=element_blank()) +
    geom_density2d(aes(color=tool),bins=25, size=.5) +
    scale_y_continuous(expand=c(0, 0), limits=c(0.3, 1.01)) +
    scale_x_continuous(expand=c(0, 0), limits=c(0.7, 1.01)) +
    geom_point(data=pt_esti,aes(specificity, sensitivity), size=3) +
    ##xlim(c(0.15, 1)) +
    ##ylim(c(0.15, 1)) +
    ##facet_wrap(~population, ncol=1) +
    theme(strip.background=element_blank(),
          legend.position=c(0.2, 0.25),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.x=element_text(size=15),
          axis.text.y=element_text(size=15),
          axis.title.x=element_text(size=20),
          axis.title.y=element_text(size=20)

          ) +
    guides(color=guide_legend(title="",
                              override.aes=list(size=1.5))) +
    xlab("Specificity") +
    ylab("Sensitivity")+theme(legend.position="none")+
  scale_color_manual(values=c("steelblue","steelblue1","lightpink1","deeppink","darkmagenta"),
                     labels=c("HE4 - 70 pM - Jacob et al., 2011","HE4 - 70 pM - JHU","CA125 - 30 U/mL - Menon et al., 2015","CA125 - 30 U/mL - JHU", "DELFI-Pro - >99% Specificity"))+theme(axis.title = element_text(face = "bold"),
        axis.text.x=element_text(face = "bold"), 
        axis.text.y=element_text(face = "bold"))
panelA2
#panelA2 <- ggplotGrob(panelA2)
```

```{r contours, fig.width=15, fig.height=15, dev=c("png", "pdf")}

set1<-panelA+theme(legend.position="none")#+patchwork::inset_element(panelA2,.1,.1,.6,.6)

l<-get_legend(panelA)

set1

l

panelA2

```


#I CHANGED THIS SECTION -- ASK ROB

```{r prevalence, fig.width=10, fig.height=6}
#prev.wide <- rdirichlet(1000, cohorts$prevalence*1000) %>%
#    as_tibble() %>%
#    set_colnames(cohorts$disease)

#prev <- epi.betabuster(mode=cohorts$prevalence[1],
#                             conf=0.975,
#                             greaterthan=TRUE,
#                             x=cohorts$prevalence[1]/10,
#                             max.shape1=500,
#                             conf.level=0.95)
s1<-236511
#s1<-19680
#s1<-2055
#s2<-(274633642/2)-s1
s2<-63757324-s1
#s2<-202562-s1
#prev <- rbeta(1000, shape1=prev$shape1, shape2=prev$shape2)
prev <- rbeta(1000, shape1=s1, shape2=s2)

#prev.long <- prev.wide %>%
#    pivot_longer(cols=all_of(cohorts$disease), names_to="Disease",
#                 values_to="Prevalence")

prev.long<-tibble(Disease="Ovarian Cancer",Prevalence=prev)
prev.long %>%
    ggplot(aes(Prevalence)) +
    geom_density(color="gray", fill="gray") +
    theme_bw(base_size=15) +
    facet_wrap(~Disease, scales="free_y") +
    theme(panel.grid=element_blank()) +
    xlim(c(0, 1))
```

```{r update_params}
prev3 <- tibble(prevalence=prev)
params4$prevalence <- rep(prev3, 5)
```

## Simulations

1. sample 100,000 people from high risk population (multinomial)

2. subsample based on adherence

3. calculate performance statistics 

4. repeat 1-3 1000 times

```{r functions}
perf <- function(i, object, N=100e3){
    dat <- unnest(object[i, ], c("performance",
                                 "adherence",
                                 "prevalence"))
    L <- nrow(dat)
    set.seed(149491) ## use same random number seed to capture differences in parameters
    screened <- rbinom(L, size=N, prob=dat$adherence)
    ## prevalence
    P <- rbinom(L, size=screened, prob=dat$prevalence) ## P = FN + TP
    N <- screened - P  ## N = TN + FP
    TP <- rbinom(L, size=P, prob=dat$sensitivity)
    FP <- rbinom(L, size=N, prob=(1-dat$specificity))
    TN <- N - FP  ## N = FP + TN
    FN <- P - TP
    fpr <- FP/N
    fnr <- FN/P
    tnr <- TN/N
    tpr <- TP/P ## sensitivity
    acc <- (TP+TN)/(P+N)
    err <- (FP+FN)/(P+N)
    ppv <- TP/(TP+FP)
    npv <- TN/(TN+FN)
    stats <- tibble("P"=P, "N"=N, "TP"=TP, "FP"=FP, "TN"=TN,
                    "FN"=FN, "acc"=acc, "err"=err,
                    "fpr"=fpr, "fnr"=fnr, "tnr"=tnr,
                    "tpr"=tpr, "ppv"=ppv, "npv"=npv,
                    "number_screened"=screened)
    stats2 <- bind_cols(dat, stats)
    stats2
}
```

```{r simulation}
## for now, we only care about HCC versus not HCC
params5 <- seq_len(nrow(params4)) %>%
    map_dfr(perf, params4) #%>%
    #filter(tool %in% c("DELFI", "US + AFP"))
#There are 10 simulations in which there were 0 cases of ovarian cancer, so FNR and TPR end up NaN but should be 0
params5 <- params5 %>% mutate(fnr=if_else(is.na(fnr),0,fnr))
params5 <- params5 %>% mutate(tpr=if_else(is.na(tpr),0,tpr))

montecarlo <- params5



```

I want to plot all metrics
```{r fpr, fig.width=8, fig.height=15, dev=c("png", "pdf")}

dat<-montecarlo %>% select(fpr,tool)

dat<-dat %>% gather(key=metric,value=score,-tool)

#dat<-dat %>% filter(tool %in% c("CA125_30_ukctocs","CA125_30_jhu","HE4_70_Jacob","HE4_70_jhu","DELFI-Pro"))
dat$tool<-factor(dat$tool,levels=c("HE4_70_Jacob","HE4_70_jhu","CA125_30_ukctocs","CA125_30_jhu","DELFI-Pro"))

#colors=c("goldenrod1","darkorange","goldenrod3","darkorange2","goldenrod4","darkorange3","dodgerblue","cyan2","cyan4","violetred4")

set2<-ggplot(dat,aes(x=tool,y=score,fill=tool))+geom_boxplot(outlier.shape=NA)+theme_classic(base_size=20)+ylim(0,.5)+
  scale_fill_manual(values=c("steelblue","steelblue1","lightpink1","deeppink","darkmagenta"),
                     labels=c("HE4 - 70 pM - Jacob et al., 2011","HE4 - 70 pM - JHU","CA125 - 30 U/mL - Menon et al., 2015","CA125 - 30 U/mL - JHU", "DELFI-Pro - >99% Specificity"))+theme(axis.text.x=element_text(angle=60,vjust = 1, hjust=1))+xlab("")+ylab("Predicted FPR")+theme(legend.position="none")+
  scale_x_discrete(labels=c("HE4-Jacob","HE4-JHU","CA125-Menon","CA125-JHU","DELFI-Pro"))+theme(axis.title = element_text(face = "bold"),
        axis.text.x=element_text(face = "bold"), 
        axis.text.y=element_text(face = "bold"))
set2

dat %>% group_by(tool) %>% summarize(mean(score))
dat %>% group_by(tool) %>% summarize(med=median(score),perc_fpr=med*100,ci_low=round(quantile(score, 0.025), 4)*100,ci_high=round(quantile(score, 0.975), 4)*100)#,sd=sd(score)*100)

```

```{r ppv, fig.width=8, fig.height=15, dev=c("png", "pdf")}

dat<-montecarlo %>% select(ppv,tool)

dat<-dat %>% gather(key=metric,value=score,-tool)

#dat<-dat %>% filter(tool %in% c("CA125_30_ukctocs","CA125_30_jhu","HE4_70_Jacob","HE4_70_jhu","DELFI-Pro"))
dat$tool<-factor(dat$tool,levels=c("HE4_70_Jacob","HE4_70_jhu","CA125_30_ukctocs","CA125_30_jhu","DELFI-Pro"))

#colors=c("goldenrod1","darkorange","goldenrod3","darkorange2","goldenrod4","darkorange3","dodgerblue","cyan2","cyan4","violetred4")

set3<-ggplot(dat,aes(x=tool,y=score,fill=tool))+geom_boxplot(outlier.shape=NA)+theme_classic(base_size=20)+ylim(0,.9)+
  scale_fill_manual(values=c("steelblue","steelblue1","lightpink1","deeppink","darkmagenta"),
                     labels=c("HE4 - 70 pM - Jacob et al., 2011","HE4 - 70 pM - JHU","CA125 - 30 U/mL - Menon et al., 2015","CA125 - 30 U/mL - JHU", "DELFI-Pro - >99% Specificity"))+theme(axis.text.x=element_text(angle=60,vjust = 1, hjust=1))+xlab("")+ylab("Predicted PPV")+theme(legend.position="none")+
    scale_x_discrete(labels=c("HE4-Jacob","HE4-JHU","CA125-Menon","CA125-JHU","DELFI-Pro"))+theme(axis.title = element_text(face = "bold"),
        axis.text.x=element_text(face = "bold"), 
        axis.text.y=element_text(face = "bold"))


set3

dat %>% group_by(tool) %>% summarize(mean(score))
dat %>% group_by(tool) %>% summarize(med=median(score),perc_ppv=med*100,ci_low=round(quantile(score, 0.025), 4)*100,ci_high=round(quantile(score, 0.975), 4)*100)#,sd=sd(score)*100)

```

```{r full,fig.width=20, fig.height=8, dev=c("png", "pdf")}
plot_grid(set1,set3,set2,ncol=3,rel_widths = c(1,.75,.75),align="h",axis="b")
```



```{r stats_adherence}
#mc <- filter(montecarlo, population=="Early stage",
#             tool %in% c("DELFI", "US + AFP"))
mc<-montecarlo

ppv <- mc %>%
    group_by(tool) %>%
    summarize(mu=mean(ppv),
              `0.025`=quantile(ppv, prob=0.025),
              `0.975`=quantile(ppv, prob=0.975)) %>%
    mutate_at(2:4, round, 3) %>%
    mutate_at(2:4, scales::percent) %>%
    set_colnames(c("tool", "mu", "q2.5", "q97.5")) %>%
    unite("ci", c("q2.5", "q97.5"), sep="-")

fpr <- mc %>%
    group_by(tool) %>%
    summarize(mu=mean(fpr),
              `0.025`=quantile(fpr, prob=0.025),
              `0.975`=quantile(fpr, prob=0.975)) %>%
    mutate_at(2:4, round, 3) %>%
    mutate_at(2:4, scales::percent) %>%
    set_colnames(c("tool", "mu", "q2.5", "q97.5")) %>%
    unite("ci", c("q2.5", "q97.5"), sep="-")

ppv

fpr
```



