---
title: "Fig 4"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---


```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(RColorBrewer)
library(grid)
library(gridExtra)
library(here)
library(ggstance)
library(readxl)
library(devtools)
devtools::load_all(here("code","liver.tools"))
```
Read in Noushin's data for TCGA figure

```{r data}
fig.data <- readRDS(here('data', 'tcga', 'data.rds'))
```

## TCGA copy number

TCGA figure panel -- problems, chromosome widths, making the panel match size of one facet in z-score panel

```{r plot, fig.height= 23, fig.width=8,dev=c("png", "pdf")}
setDT(fig.data)
fig.data[,bin:=as.factor(rev(bin))][]
fig.data$disease<-"TCGA Ovarian Cancers\n(n = 597)"
plot <- fig.data %>%
    mutate(disease="TCGA Ovarian Cancers\n(n = 597)") %>%
    ggplot(aes(x=value, y=bin, color=change, fill=change)) +
    facet_grid(arm~disease, scales = 'free_y') + ##,
    ##           switch = 'y')+#, space = 'free')+
    ##facet_wrap(~arm, ncol=1, scales="free_y") + ##
    geom_col() +
    scale_x_continuous() +
    #scale_fill_brewer(palette = 'Set1')+
    #scale_color_brewer(palette = 'Set1')+
    scale_color_manual(values=c("red3","#18468B"))+
    scale_fill_manual(values=c("red3","#18468B"))+
    theme_minimal(base_size=24) +
    theme(strip.text.y = element_text(angle = 0,
                                      hjust = 0.5,
                                      vjust = 0.5,
                                      size=20),
          strip.text.x=element_text(size=20),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          panel.grid =  element_blank(),
          legend.title = element_blank(),
          legend.position="bottom",
          axis.title.x = element_blank(),
          strip.background = element_blank()) + ##removing label
    #xlab("Proportion of\ncases with CNV") +
    ylab("Chromosome Arm")
```

## Ovarian study z-scores

The z-score's with our data panel. Problems -- chromosome width.

```{r z_scores}
meta<-fread(here("data","metadata.csv"))
meta<-meta %>% select(-V1) %>% mutate(classifier_type2=if_else(id=="CGPLOV161P","Benign",classifier_type2))

data<-fread(here("data","screening_train_data.csv")) %>% select(-type)
data2<-fread(here("data","diag_train_data.csv")) %>% filter(type=="healthy") %>% select(-type)
data<-rbind(data,data2)

data<-inner_join(data,meta %>% select(id,classifier_type),by="id") %>% rename(type=classifier_type)



features<-data

cols <- brewer.pal(7,"Set1")
cols[3] <- cols[4]

features <- features %>% select(id,starts_with("z"))
features <- inner_join(features, meta %>%
                                     select(id,`classifier_type2`),
                           by=c("id"="id"))




datx<-read_csv(here("data/scores","All_Scores_CV.csv"))
datx<-datx %>% filter(training_set_include_screening=="yes"|training_set_include_diag=="yes")
datx<-datx %>% select(id,classifier_type,`Screening ZEUS`,`Diagnostic ZEUS`,subtype_simple,stage)
datx<-datx %>% gather(key=model,value=score,-stage,-subtype_simple,-classifier_type,-id) %>% filter(!is.na(score))

features<-features %>% filter(id %in% datx$id)

inner_join(features %>% select(id,classifier_type2),datx %>% select(model,id,classifier_type),by="id")

features <- features %>%
        gather(key="Arm", value="Zscore", starts_with("z"))
    features$Arm <- sapply( str_split(features$Arm,"_"), "[", 2 )

    chrlevels <- c("1p","1q","2p","2q","3p","3q","4p","4q","5p",
                   "5q","6p","6q","7p","7q","8p","8q","9p","9q",
                   "10p","10q","11p","11q","12p","12q",
                   "13q","14q","15q","16p","16q","17p","17q",
                   "18p","18q","19p","19q","20p","20q","21q","22q")
    features$Arm <- factor(features$Arm,
                           levels = chrlevels)
    features <- features %>% mutate(colors=if_else(Zscore<=0,"Negative","Positive"))
    ##features$colors<-factor(features$colors,levels=c("Positive","Negative"))
    ##Z-score for two-tailed p-value .05 w/ bonferroni correction for 39 arms
    features <- features %>%
        mutate(colors=if_else(Zscore>= -3.22 & Zscore <= 3.22 ,
                              "Neutral", features$colors))
    ##features<- features %>% mutate(transp=if_else(colors=="Neutral",.9,.2))
    ##features<- features %>% mutate(transp=if_else(colors=="Neutral",.05,.99))
    features <- features %>% mutate(transp=if_else(colors=="Neutral",1,1))
    ##features<-features %>% mutate(colors=if_else(Zscore<=0,"Negative","Positive"))
    features$colors<-factor(features$colors,levels=c("Positive","Negative","Neutral"))
    features$y=1
    ##features$log<-log10(abs(features$Zscore))
    ##features <- features %>% mutate(log=if_else(Zscore<=0,-1*log,log))
    features <- features %>% mutate(Zscore=if_else(Zscore<= -75,-75,Zscore))
    features <- features %>% mutate(Zscore=if_else(Zscore>75,75,Zscore))
    features <- features %>%
        mutate(`type`=if_else(`classifier_type2`=="cancer","Ovarian Cancer",classifier_type2))
      features <- features %>%
        mutate(`type`=if_else(`classifier_type2`=="non-cancer","Non-cancer",type))

    features$type <- factor(features$`type`,
                                  levels=c("Ovarian Cancer","Benign","Non-cancer"))
    features2  <- features %>%
        rename(z=Zscore) %>%
        mutate(root2=nroot(z, 2)) %>%
        mutate(root3=nroot(z, 3))


features2 <- features2  %>%
    select(id, type, Arm,
           z, root2, root3, y, colors, transp)
###YOU ARE HERE AFTER LUNCH!!!
ns <- features2 %>%
    group_by(type) %>%
    summarize(n=length(unique(id))) %>%
    mutate(label=paste0(type, "\n(n = ", n, ")"))
ns <- setNames(ns$label, ns$type)
cols[2]<-"#18468B"
cols[1]<-"red3"

zplot_cancer <- features2 %>%
    ggplot(aes(x=y, y=z, color=colors, alpha=transp)) +
    geom_point(size=5) +
    facet_grid(Arm ~ type, scales = 'free_y',labeller=labeller(type=ns)) +
    coord_flip() +
    theme_minimal(base_size=24) +
    theme(strip.text.y.left = element_text(angle = 0,
                                           hjust = 0.5,
                                           vjust = 0.5),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),

          panel.grid =  element_blank(),
          plot.title = element_blank(),
          legend.title = element_blank(),
          axis.title.y=element_blank(),
          legend.position = "bottom",
          strip.text.y = element_blank(),
          axis.title.x = element_blank()) + #####removing label
    scale_y_continuous(limits = c(-75,75),breaks=c(-70,0,70)) +
    scale_alpha_identity() +
    scale_fill_manual(values = cols) +
    scale_color_manual(values = cols) #+
    #ylab("Z-Score")
```


```{r z_scores2}
meta<-fread(here("data","metadata.csv"))
meta<-meta %>% select(-V1) %>% mutate(classifier_type2=if_else(id=="CGPLOV161P","Benign",classifier_type2))

data<-fread(here("data","screening_test_data.csv")) %>% select(-type)
data2<-fread(here("data","diag_test_data.csv")) %>% filter(type=="healthy") %>% select(-type)
data<-rbind(data,data2)

data<-inner_join(data,meta %>% select(id,classifier_type),by="id") %>% rename(type=classifier_type)



features<-data

cols <- brewer.pal(7,"Set1")
cols[3] <- cols[4]

features <- features %>% select(id,starts_with("z"))
features <- inner_join(features, meta %>%
                                     select(id,`classifier_type2`),
                           by=c("id"="id"))




datx<-read_csv(here("data/scores","All_Scores_Val.csv"))
datx<-datx %>% filter(screen_val=="yes"|diag_val=="yes")
datx<-datx %>% select(id,classifier_type,`Screening ZEUS`,`Diagnostic ZEUS`,subtype_simple,stage)
datx<-datx %>% gather(key=model,value=score,-stage,-subtype_simple,-classifier_type,-id) %>% filter(!is.na(score))

features<-features %>% filter(id %in% datx$id)

features <- features %>%
        gather(key="Arm", value="Zscore", starts_with("z"))
    features$Arm <- sapply( str_split(features$Arm,"_"), "[", 2 )

    chrlevels <- c("1p","1q","2p","2q","3p","3q","4p","4q","5p",
                   "5q","6p","6q","7p","7q","8p","8q","9p","9q",
                   "10p","10q","11p","11q","12p","12q",
                   "13q","14q","15q","16p","16q","17p","17q",
                   "18p","18q","19p","19q","20p","20q","21q","22q")
    features$Arm <- factor(features$Arm,
                           levels = chrlevels)
    features <- features %>% mutate(colors=if_else(Zscore<=0,"Negative","Positive"))
    ##features$colors<-factor(features$colors,levels=c("Positive","Negative"))
    ##Z-score for two-tailed p-value .05 w/ bonferroni correction for 39 arms
    features <- features %>%
        mutate(colors=if_else(Zscore>= -3.22 & Zscore <= 3.22 ,
                              "Neutral", features$colors))
    ##features<- features %>% mutate(transp=if_else(colors=="Neutral",.9,.2))
    ##features<- features %>% mutate(transp=if_else(colors=="Neutral",.05,.99))
    features <- features %>% mutate(transp=if_else(colors=="Neutral",1,1))
    ##features<-features %>% mutate(colors=if_else(Zscore<=0,"Negative","Positive"))
    features$colors<-factor(features$colors,levels=c("Positive","Negative","Neutral"))
    features$y=1
    ##features$log<-log10(abs(features$Zscore))
    ##features <- features %>% mutate(log=if_else(Zscore<=0,-1*log,log))
    features <- features %>% mutate(Zscore=if_else(Zscore<= -75,-75,Zscore))
    features <- features %>% mutate(Zscore=if_else(Zscore>75,75,Zscore))
    features <- features %>%
        mutate(`type`=if_else(`classifier_type2`=="cancer","Ovarian Cancer",classifier_type2))
      features <- features %>%
        mutate(`type`=if_else(`classifier_type2`=="non-cancer","Non-cancer",type))

    features$type <- factor(features$`type`,
                                  levels=c("Ovarian Cancer","Benign","Non-cancer"))
    features2  <- features %>%
        rename(z=Zscore) %>%
        mutate(root2=nroot(z, 2)) %>%
        mutate(root3=nroot(z, 3))


features2 <- features2  %>%
    select(id, type, Arm,
           z, root2, root3, y, colors, transp)
###YOU ARE HERE AFTER LUNCH!!!
ns <- features2 %>%
    group_by(type) %>%
    summarize(n=length(unique(id))) %>%
    mutate(label=paste0(type, "\n(n = ", n, ")"))
ns <- setNames(ns$label, ns$type)
cols[2]<-"#18468B"
cols[1]<-"red3"

zplot_cancer_val <- features2 %>%
    ggplot(aes(x=y, y=z, color=colors, alpha=transp)) +
    geom_point(size=5) +
    facet_grid(Arm ~ type, scales = 'free_y',labeller=labeller(type=ns)) +
    coord_flip() +
    theme_minimal(base_size=24) +
    theme(strip.text.y.left = element_text(angle = 0,
                                           hjust = 0.5,
                                           vjust = 0.5),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),

          panel.grid =  element_blank(),
          plot.title = element_blank(),
          legend.title = element_blank(),
          axis.title.y=element_blank(),
          legend.position = "bottom",
          strip.text.y = element_blank(),
          axis.title.x = element_blank()) + #####removing label
    scale_y_continuous(limits = c(-75,75),breaks=c(-70,0,70)) +
    scale_alpha_identity() +
    scale_fill_manual(values = cols) +
    scale_color_manual(values = cols) #+
    #ylab("Z-Score")
```
Adding together the TCGA and z-score panel -- problems -- chromosome width, size of TCGA panel, not yet stacked with feature importance panel.

```{r cowplot, fig.height=30, fig.width=25,dev=c("png", "pdf","svg")}
#pdf("test.pdf")
cowplot::plot_grid(plot,zplot_cancer,zplot_cancer_val,nrow=1,align="v",axis="l")
#dev.off() #
```

